### 2025-01-23 BAI/exercises.tex
Added lecture questions.

% Visual information processed in different brain areas
%  Images projected from retina to striate and extrastriate cortex,
% filtered according to different features (edges, orientation, colour,
% motion, …)
%  Visual Processing inspired
% hierarchical computational models for:
% • Object recognition (Fukushima, LeCun)
% • Deep Learning models
\begin{exercise}{}
  % TODO

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

% Models for bio-inspired sound source localization, visual
% localization and multisensory integration
%  Multisensory VR environment for neuro-robotic experiments
%  Effectiveness of bio-inspired models
%  Demonstrated behavioral, neurophysiological
% similarity between models and biological systems
\begin{exercise}{}
  % TODO

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

% Advantages:
%  Fast Reaction – Immediate mapping of sensory information onto motor actions
%  Robustness – If one part fails, robot may retain some behavioral competence
%  Multiple Goals – Can follow multiple goals simultaneously by combining
% behaviors
%  Extensibility – Easy to add new parts on top of previous
%  Simplicity – Complexity derives from continuous
% interaction of simple modules with environment
% and each other
%  Computational tractability – Usually simple calculations
% and easy to parallelize
%  Development of using LLM for planning
% Disadvantages/Limits:
%  Limited Information – Agents without environment models must have
% sufficient information from local environment
%  Non-Local information – how does agent take into account non-local
% information (e.g. memory?)
%  Learning globally – Difficult to make reactive agents
% that learn global behavior
%  Complex dynamics – It is hard to engineer agents
% with large numbers of behaviors (dynamics of
% interactions become too complex to understand)
\begin{exercise}{}
  % TODO

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

% Gestures are an integral part in communication
% • Important in multimodal systems
% • Research on autonomous robots and HRI
%  Model alternatives and choice
% • Dependent on gesture type → gesture continuum
% • Deep networks provide partial solution for feature extraction
% • Recurrent networks in LSTM and GammaGWR
% • Reservoir Computing models by neurophysiological principles
% for action selection in prefrontal cortex and striatum
% • Tradeoff between benchmarking and developing cognitive systems
\begin{exercise}{}
  % TODO

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

% Pros:
% • General heuristic for problem search
% • Able to find good solutions in feasible computing time
% • Distributed execution possible
%  Cons:
% • Cannot guarantee optimal solutions
% • Long runtimes, compared to search algorithms
% • Success depends also on used parameters
%  Many further EA approaches: NEAT, HyperNEAT…
\begin{exercise}{}
  % TODO

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

%  LMM deliver pointwise understanding for frequent cases
%  For rare cases LMM often “autocorrect reality” answering often wrong question
%  Hybrid systems make complementary use of symbolic and connectionist AI
%  Symbolic rules can be extracted from neural networks
%  Different architectures for integrating symbolic and neural computing are
% possible
%  How images are processed by a neural network can be visualized
%  Hybrid systems and explainable AI are in the
% current focus of research
%  Topics for thesis, independent studies, etc.
\begin{exercise}{}
  % TODO

  \begin{solution}
    % TODO
  \end{solution}
\end{exercise}

@@

\begin{exercise}{Multiple Choice}
  \begin{enumerate}
    \item In which models is backpropagation used for learning?
      \begin{itemize}
        \item Multilayer perceptron (MLP)
        \item Convolutional neural network (CNN)
        \item Self-organizing map (SOM)
        \item None of the above
      \end{itemize}

    \item What does "embodied" mean for language?
      \begin{itemize}
        \item Language is based on the physical world.
        \item Language is based on the ground truth.
        \item Language is based on the ground state.
        \item Language is based on the ground floor.
      \end{itemize}

    \item What are the two streams and their tasks in human vision?
      \begin{itemize}
        \item Dorsal stream for spatial vision and ventral stream for object recognition
        \item Ventral stream for spatial vision and dorsal stream for object recognition
        \item Dorsal stream for object recognition and ventral stream for spatial vision
        \item Ventral stream for object recognition and dorsal stream for spatial vision
      \end{itemize}
  \end{enumerate}

>>>>

\begin{exercise}{Hierarchical Vision}
  \begin{enumerate}
    \item Which visual information is processed in which brain areas?
    \item What are the features filtered in the visual processing of the brain?
    \item Explain a hierarchical computational model for object recognition.
    \item How can deep learning be utilized for visual processing?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item Low-level visual information is processed in the primary visual cortex, while high-level visual information is processed in the ventral and dorsal streams of the brain.
      \item The features filtered in the visual processing of the brain include edges, textures, shapes, and objects.
      \item A hierarchical computational model for object recognition involves multiple layers of processing, where each layer extracts increasingly complex features from the input data to recognize objects.
      \item Deep learning can be utilized for visual processing by using deep neural networks to learn hierarchical representations of visual data and perform tasks like image recognition and object detection.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Crossmodal Processing}
  \begin{enumerate}
    \item Briefly explain one model for bio-inspired sound source localization, visual localization, and multisensory integration each.
    \item How effective are bio-inspired models in practice?
    \item Demonstrate similarities and differences between bio-inspired models and biological systems.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item One model for bio-inspired sound source localization is the duplex theory of sound localization, which uses interaural time difference (ITD) and interaural level difference (ILD) to localize sound sources. One model for bio-inspired visual localization is the hierarchical processing model, which uses multiple layers of processing to extract features from visual data. One model for bio-inspired multisensory integration is the Bayesian integration model, which combines information from multiple sensory modalities to make decisions.
      \item Bio-inspired models are effective in practice for tasks like sound source localization, visual localization, and multisensory integration. These models can capture the complexity and flexibility of biological systems and provide insights into how sensory information is processed in the brain.
      \item Bio-inspired models demonstrate behavioral and neurophysiological similarities with biological systems, such as the ability to localize sound sources, recognize objects, and integrate information from multiple sensory modalities. However, there are also differences between bio-inspired models and biological systems, such as the level of abstraction and complexity of the models.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Behavior-based Robotics}
  \begin{enumerate}
    \item Explain the advantages and disadvantages of behavior-based robotics.
    \item How can behavior-based robotics be used for planning?
    \item What are the limits of behavior-based robotics?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item The advantages of behavior-based robotics include fast reaction times, robustness, the ability to follow multiple goals simultaneously, extensibility, simplicity, and computational tractability. The disadvantages of behavior-based robotics include limited information, non-local information, learning globally, and complex dynamics.
      \item Behavior-based robotics can be used for planning by combining simple behaviors to achieve complex goals. This approach allows robots to react to their environment in real-time and adapt to changing conditions.
      \item The limits of behavior-based robotics include the need for sufficient information from the local environment, the challenge of incorporating non-local information, the difficulty of learning global behavior, and the complexity of interactions in large-scale systems.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Gesture recognition}
  \begin{enumerate}
    \item Explain the importance of gestures in communication and their role in multimodal systems.
    \item Which models are suitable for gesture recognition and how do they differ?
    \item How would you choose a model for gesture recognition?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item Gestures are an integral part of communication and play an important role in multimodal systems. They provide additional information beyond speech and can convey emotions, intentions, and context.
      \item Models suitable for gesture recognition include deep networks for feature extraction, recurrent networks like LSTM and GammaGWR for sequence modeling, and reservoir computing models for action selection. These models differ in their approach to feature extraction, sequence modeling, and action selection.
      \item When choosing a model for gesture recognition, it is important to consider the type of gestures being recognized, the complexity of the task, the amount of data available, and the computational resources required. Deep networks are suitable for feature extraction, recurrent networks are suitable for sequence modeling, and reservoir computing models are suitable for action selection.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Evolutionary Computing}
  \begin{enumerate}
    \item Explain the advantages and disadvantages of evolutionary algorithms.
    \item What are the limits of evolutionary algorithms?
    \item Name and describe two evolutionary algorithms.
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item The advantages of evolutionary algorithms include being a general heuristic for problem search, the ability to find good solutions in feasible computing time, and the possibility of distributed execution. The disadvantages of evolutionary algorithms include not guaranteeing optimal solutions, long runtimes compared to search algorithms, and the success depending on the parameters used.
      \item The limits of evolutionary algorithms include the need for a large number of evaluations, the difficulty of tuning parameters, the risk of premature convergence, and the challenge of scaling to large problems.
      \item Two evolutionary algorithms are the genetic algorithm, which uses selection, crossover, and mutation operators to evolve a population of solutions, and the genetic programming algorithm, which evolves programs or models to solve a problem.
    \end{enumerate}
  \end{solution}
\end{exercise}

\begin{exercise}{Neuro-Symbolic and Explainable AI Systems}
  \begin{enumerate}
    \item Explain the concept of neuro-symbolic systems and how they combine symbolic and connectionist AI.
    \item How can symbolic rules be extracted from neural networks?
    \item Describe different architectures for integrating symbolic and neural computing.
    \item How can images be processed by a neural network and visualized?
    \item Why are hybrid systems and explainable AI in the current focus of research?
  \end{enumerate}

  \begin{solution}
    \begin{enumerate}
      \item Neuro-symbolic systems combine symbolic and connectionist AI by integrating symbolic reasoning with neural network processing. These systems use symbolic rules to guide neural network learning and decision-making.
      \item Symbolic rules can be extracted from neural networks by analyzing the learned weights and activations of the network to identify patterns and rules that explain the network's behavior.
      \item Different architectures for integrating symbolic and neural computing include neural-symbolic integration, where symbolic reasoning is combined with neural network processing, and neural-symbolic learning, where neural networks are used to learn symbolic rules from data.
      \item Images can be processed by a neural network using convolutional layers to extract features from the input data and fully connected layers to classify the features. The processing steps can be visualized by examining the activations and weights of the network at each layer.
      \item Hybrid systems and explainable AI are in the current focus of research because they provide a way to combine the strengths of symbolic and connectionist AI, improve the interpretability of AI systems, and enable more transparent and accountable decision-making.
    \end{enumerate}
  \end{solution}
\end{exercise}

@@

\begin{exercise}{Multiple Choice}
  \begin{enumerate}
    \item In 
      \begin{itemize}
        \item Multilayer perceptron (MLP)
        \item Convolutional neural network (CNN)
        \item Self-organizing map (SOM)
        \item none of the above
      \end{itemize}
        is backpropagation used for learning.

    \item An embodied language is based on
      \begin{itemize}
        \item the physical world.
        \item ground truth.
        \item ground state.
        \item ground floor.
      \end{itemize}

    \item What are the two streams and their tasks in human vision?
      \begin{itemize}
        \item Dorsal stream for spatial vision and ventral stream for object recognition
        \item Ventral stream for spatial vision and dorsal stream for object recognition
      \end{itemize}
  \end{enumerate}
